High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2020-10-20 paper/main.tex
updated description of synthetic data with given high-order dynamics

\subsubsection*{Simulated higher-order correlations}
We simulated first and second order correlations reflecting the 4 qualitatively different patterns
of synthetic dynamic correlations.
 We repeated the processes 10 times for each order and for each of the 4 dynamic pattern (Constant, Random, Ramping, and
Event) explored in \textit{Synthetic data}.  First,
 we generated a timeseries of correlations which comprised 50 features (dimensions)
that varied over 300 timepoints using 1 of the 4 \textit{Synthetic data} dynamic
patterns. To calculate the next increasing order of corrrelations, ${ Y_{n+1,t}}$, we
multiplied the vectorized correlation matrix by its transpose, for
each timepoint:

\begin{align}
Y_{n=1,t} = vec(\mathbf{\Sigma}_{t})\cdot
  vec(\mathbf{\Sigma}_{t})^\top \\
  Y_{n+1,t} = vec(Y_{n,t}) \cdot vec(Y_{n,t})^\top\label{eqn:veccorr}
  \end{align}

  We repeated this processes for the number of desired orders (N) to
  generate a timeseries of $\mathbf{N^{th}}$ correlation matrices.  Then, observations were drawn at each timepoint from a  ${ Y_{N-1,t}}$-mean multivariate Gaussian distribution with a
 covariance matrix defined as ${ Y_{N+1,t}}$ and the resultant vector
 was reshaped to the matrix the size of ${ Y_{N-1,t}}$:
 \begin{align}
   { \hat{Y}_{N-1,t}} \sim vec(\mathcal{N}({ Y_{n,t}}, { Y_{N,t}}))^{-1}
   \end{align}
 This process was then repeated for all decreasing orders, in which
 observations werre drawn at each timepoint from a  ${ Y_{n,t}}$-mean
 multivariate Gaussian distribution, but the covariance matrix is
 defined at the previously estimated reshaped matrix $\hat{Y}_{n,t}$:
 \begin{align}
   { \hat{Y}_{n,t}} \sim vec(\mathcal{N}({ Y_{n,t}}, {\hat{ Y}_{n+1,t}}))^{-1}
 \end{align}
We repeat this process until $n = 1$, then we drew observations at each timepoint
from a zero-mean multivariate Gaussian distribution with a
covariance matrix defined for each timepoint as described below:

\begin{align}
 s_t \sim \mathcal{N}(\mathbf{0},  \hat{Y}_{n=1,t})
\end{align}

For our specific analyses the resulting synthetic timeseries comprised
50 features and 300 timepoints. We then used this synthetic data to test how well Equation~\ref{eqn:timecorr}
could recover the simulated first and second order correlations.  To mitigate temporal blur in
estimating each timeseries of dynamic correlations, we used a Dirac $\delta$ function
kernel (which places all of its mass over a single timepoint)
Fig.~\ref{fig:kernels}b) to compute each lower-order correlation and then used a
Laplace kernel with a width of 20 to compute  $\mathbf{ X_{n,t}}$, applying
temporal blur at just the last step.  Then, for each timepoint, we
correlated the simulated correlation matix, ${ Y_{n,t}}$ with the
Equation~\ref{eqn:timecorr} estimated correrlation matric, $\mathbf{ X_{n,t}}$.

>>>>

\subsubsection*{Synthetic data: simulating dynamic high-order
  correlations}

We developed an iterative procedure for constructing timeseries data
that exhibits known dynamic high-order correlations.  The procedure
builds on our approach to generating dynamic first-order correlations.
Essentially, once we generate a timeseries with known first-order
correlations, we can use the known first-order correlations as a
template to generate a new timeseries of second-order correlations.
In turn, we can generate a timeseries of third-order correlations from
the second-order correlations, and so on.  In general, we can generate
order $n$ correlations given a timeseries of order $n - 1$ correlations,
for any $n > 1$.  Finally, given the order $n$ timeseries, we can
reverse the preceding process to generate an order $n - 1$ timeseries, an
order $n - 2$ order timeseries, and so on, until we obtain an order 0
timeseries of simulated data that reflects the chosen high-order dynamics.

The central mathematical operations in our procedure are two
functions, $\mathrm{vec}(\cdot)$ and $\mathrm{mat}(\cdot)$.  The
$\mathrm{vec}(\cdot)$ function takes as input a $K \times K$ symmetric
matrix and returns as output a
$\left(\frac{K^2 - K}{2} + K\right)$-dimensional column vector
containing the entries in the upper triangle and diagonal.  The
$\mathrm{mat}(\cdot)$ function inverts $\mathrm{vec}(\cdot)$ by taking
as input a $\left(\frac{K^2 - K}{2} + K\right)$-dimensional column
vector and returning a $K \times K$ symmetric matrix as output.  We
can then generate an order $n$ correlation matrix (for one timepoint, $t$) from an order
$n - 1$ template (from the same timepoint) as follows:
\begin{align}
  \mathbf{\Sigma_{n}}(t) &= \mathrm{mat}(\mathrm{vec}(\mathbf{\Sigma_{n-1}}(t)) \otimes \mathrm{vec}(\mathbf{\Sigma_{n-1}}(t))^\top).\label{eqn:veccorr}
\end{align}

Given a timeseries of order $n$ correlation matrices, we can draw an
order $n-1$ correlation matrix for each timpoint $t$ using
\begin{align}
 \mathbf{\sigma_{n-1}}(t) &\sim \mathcal{N}\left(\mathbf{0}, \mathbf{\Sigma_n}(t)
                   \right)\label{eqn:highorder-gen1}\\
  \mathbf{\Sigma_{n-1}}(t) &= \mathrm{mat}(\mathbf{\sigma_{n-1}}(t)).\label{eqn:highorder-gen2}
  \end{align}
  We can then use repeated applications of
  Equations~\ref{eqn:highorder-gen1} and~\ref{eqn:highorder-gen2} in
  order to obtain a synthetic dataset.

 When the template first-order correlations are constructed to exhibit different
 temporal profiles (e.g., using the constant, random, ramping, and
 event procedures described above), the resulting high-order
 correlations and synthetic data will exhibit the same category of
 temporal profile.

 Following our approach to generating synthetic data exhibiting known
 first-order correlations, we constructed a total of 40 additional multivariate timeseries,
collectively reflecting a total of 4 qualitatively different patterns
of dynamic correlations (i.e., 10 datasets reflecting each type of
dynamic pattern: constant, random, ramping, and event).
